\documentclass[main.tex]{subfiles}
\begin{document}
%%%%%%%%%%%%%%%%%%%%
\section{Algebra}
For algebra, the problem seems to fall into three categories: group theory, rep theory, and ring/Galois theory.

\subsection{Group Theory}
The most important thing here is the Sylow theorems and how they are used. Sylow's theorem is about subgroups of prime orders. The idea is looking at the orbits of the conjugation action of $G$ on $G$.



\begin{definition}
Let $G$ be a finite group of order $p^n m$ where $m$ is coprime to $p$ to $p$. A Sylow $p$-subgroups of $G$ is a subgroup of order $p^n$. They are the maximal $p$-subgroups.
\end{definition}

Here the Sylow theorems:

\begin{theorem}[Sylow's 1st Theorem]
If $p$ is a prime number and $p | |G|$, then there exists a Sylow $p$-subgroup of $G$.
\end{theorem}

\begin{proof}
We find the group by looking at orbits of $G$ on subset of $G$ of size $p^n$, and then finding one $w$ with orbit size coprime to $p$. Then the stabilizer of this is $H$, which as it stabilizes it must act on the elements of the set (very confusing sentence). As it acts on the elements of $w$ freely, $H | |w| = p^n$. On the other hand, $|G|/|H| = orb(w)$ has size coprime to $p$, this means that $|H| = p^n$. (In fact $w = H$ as sets.
\end{proof}

\begin{theorem}[Sylow's 2nd Theorem]
For a given prime $p$, all Sylow $p$-subgroups are conjugate to each other.
\end{theorem}

\begin{corollary}
A Sylow $p$-subgroup of $G$ is unique iff it is normal in $G$. Thus it is unique iff it is abelian.
\end{corollary}

\begin{theorem}[Sylow's 3rd Theorem]
Let the number of Sylow $p$-subgroups to be $n_p$. Then the floowing results hold:
\begin{enumerate}
    \item $n_p = 1\ mod\ p$
    \item If $|G| = p^n m$ so that $m$ and $p$ are coprime, then $n_p | m$
    \item If $P$ is any Sylow $p$-subgroup of $G$, then $n_p = [G: N(P)]$, where $N(P)$ is the normalizer of $P$ in $G$
\end{enumerate}


\end{theorem}
\begin{remark}
part 2 and 3 of Sylow's 3rd theorem follows direct from  the facts that $G$ acts on the set of Sylow $p$-groups by conjugation, Sylow's 2nd theorem says that the action is transitive, and the stabilizer of this action at $P$ is exactly $N(P)$.
\end{remark}
\begin{corollary}
Let $G$ be a fintie group and $p$ a prime number. If $p^n$ divides the order of $G$, then $G$ has a subgroup of order $p^n$.
\end{corollary}

We also have the Cauchy's theorem:

\begin{theorem}[Cauchy's theorem]
If prime $q$ divides $|G|$, then there exists an element of order $q$.
\end{theorem}

Let $G$ acts on $G$ by conjugation, then the orbits are conjugacy classes. 
Then the class equations says that $|G| = |Z(G)| + k_1 + k_2 ..$, where $|Z(G)|$ is the center of $G$ (those that commutes with all elements those belong to a single orbit), and $k_1$ are the size of distinct non-singlet orbits.


\subsubsection{Groups of order $p^2$}
Groups of order $p$ are isomorphic to $\mathbb{Z}/p\mathbb{Z}$, namely abelian. Similar thing is true for groups of order $p^2$, they are all abelian, so either $\mathbb{Z}/p\mathbb{Z} \sum \mathbb{Z}/p\mathbb{Z}$ or $\mathbb{Z}/p^2$

\subsubsection{Groups of order $p^3$}
There are ofc abelian once $3, (2,1), (1,1,1)$, and there are two non-isomorphic nonabelian. 

\subsubsection{Groups of order $pq$}
This uses semidirect products: 
\begin{definition}
Given $N, K$ groups and an action $K$ on $N$ ($K \rightarrow Aut(N)$), then the semidirect product of $N$ and $K$ $N \rtimes K$ as set is $N \times K$, with multiplication:
$$
(a_1, b_1) (a_2, b_2) \coloneqq (a_1 (b_1 \dot a_2), b_1 b_2)
$$
\end{definition}

We have inclusion $K \rightarrow N \rtimes K$ and projection $N \rtimes K \rightarrow K$. Conversely,
Splitting of groups gives action of $N$ on the kernel of the projection, and exhibits the total group to be the semidirect sum.

\begin{proposition}
Let $p \geq q$, if $q \nmid (p-1)$ then there are only one iso class of groups of order $pq$, namely the abelian group $\mathbb{Z}/pq\mathbb{Z}$. If $q | (p-1)$, then there are two groups of order $pq$:
\begin{enumerate}
    \item The abelian one: $\mathbb{Z}/pq\mathbb{Z}$
    \item The nonabelian one given by any non-trivial 
    $$
    \phi: \mathbb{Z}/q\mathbb{Z} \rightarrow Aut(\mathbb{Z}/p\mathbb{Z}) \cong \mathbb{Z}/(p-1)\mathbb{Z}
    $$
\end{enumerate}
\end{proposition}

\begin{corollary}
For any odd prime $p$ there are two non-isomorphic groups of order $2p$:
\begin{enumerate}
    \item The cyclic group $\mathbb{Z}/2p\mathbb{Z}$
    \item The dihedral group $D_p = \mathbb{Z}/p\mathbb{Z} \rtimes \mathbb{Z}/2\mathbb{Z}$ with the sign action.
\end{enumerate}
\end{corollary}




\subsubsection{Dihedral Groups}

This leads us right into the dihedral group $D_n$. 
\begin{definition}

For any $n \geq 1$, we have the Dihedral group 
$$
D_n \coloneqq <r, s| r^n = 1, s^2 = 1, s r s = r^{-1} >
$$
This is the group of symmetry of the regular $n$ polygon, with $r$ being the a mininum rotation and $s$ being a reflection.

The group of rotation is normal, and all reflections are conjugate when $n$ is odd and there are two conjugacy classes when $n$ is even (easy to see there are two different kinds of reflections)
\end{definition}

\subsubsection{$GL_n(\mathbb{F}_p)$}
The size of finite group $GL_n(\mathbb{F}_p)$ is
$$
(p^n -1) (p^n - p) ...
$$
as this is number of choosing a nonzero vector, then choosing another vector away from the first one (the size of a $\mathbb{F}_p$ line is $p$)...

\subsection{Representation Theory}
This concerns $\mathbb{C}$-linear representation theory of finite groups, via the character theory.


\subsubsection{Basics}
Let $G$ be a finite group.
The category of (complex) finite dimensional representation $Rep(G)$ is quite simple, literally, it is a simple category:

\begin{theorem}[Maschke's Theorem]

Any short exact sequences of representations splits, and irreducibles are simples. 

\end{theorem}

One can show this by the following proposition:
\begin{proposition}
For any $G$-rep $V$ there is a $G$-equivariant Hermitian structure on $V$.
\end{proposition}
This is done by choosing any hermitian structure and then add its $G$-orbits.


The maps between irreducible representations are also simple:

\begin{theorem} [Schur's Lemma]
let $V$ and $W$ be irreducible representations, and $f : V \rightarrow W$, then $f$ is either an isomorphism or 0. Moreover, endormorphism of a irreducible representation $V$ is scalar multiplication by $\mathbb{C}$.
\end{theorem}

\begin{proof}
As $V$ and $W$ are irreducible, then the kernel of $f$ is either $V$ or 0, similiarly, the cokernel is then $0$ or $W$.

As for the second part, given $f : V \rightarrow V$, then it must have one eigenvector $v$ with eigenvalue $\lambda$ (this is why we are in $\mathbb{C}$ and not in say $\mathbb{R}$, think rotation), then $f - \lambda Id$ has a nonzero kernel, thus it is $V$. So $f = \lambda Id$.
\end{proof}

The most important representation is the regular representation, that is, $G$ act on $\mathbb{C}[G]$ by left translation.

Here's the decomposition theorem, which we can easily prove using character theory:
\begin{theorem}
There are finitely many non-isomorphic irreducible representations, in fact there are the number of conjugacy classes of $G$ many. Moreover:
$$
\mathbb{C}[G] = \bigoplus_i V_i ^{dim V_i}
$$
where $V_i$ are the distinct irreducible representations.


\end{theorem}

Next we need the character theory:
\subsubsection{Character Theory}
This is the crux of the theory of finite dimensional reps. First we define characters:

\begin{definition}
Let $\rho$ be a $G$ rep on $V$. Then the character associated to $\rho$ is a complex-valued map on the set $\frac{G}{G}$ of conjugacy classes of $G$:
$$
\chi_V(g) \coloneqq tr(\rho(g))
$$
\end{definition}
Note that $\chi_V(1) = dim V$. It turns out they encode all the info about a rep $V$. 

As $tr$ is conjugate-invariant, this is well-defined on conjugacy classes. A complex-valued map on $\frac{G}{G}$ is called a class function on $G$.

There is an inner product on class functions:
$$
(f_1, f_2) \coloneqq \frac{1}{|G|}\sum_{g \in G} f_1(g) \overline{f_2(g)}
$$

It turns out the characters of irreducible representations are orthogonal to this decomposition:

\begin{lemma}
Let $V$ and $V'$ be irreducible representations, if $V \cong V'$, then 
$$
(\chi_V, \chi_{V'}) = \delta_{V, V'}
$$

\end{lemma}

We also have another orthogonal condition:
\begin{lemma}
Given $g, h$, 
$$
\frac{1}{|G|} \sum_{V_i}\chi_{V_i}(g) \overline{\chi_{V_i}(h)} = \delta_{[g], [h]}
$$
\end{lemma}

Moreover, we have the delta functions $\delta_{[g]}$ on conjugacy classes, then we have 
$$
(\delta_{[g]}, \delta_{[h]}) = \delta_{g,h}
$$
and $\chi_{V_i}(g) = (\chi, g)$ tells us that $\chi_{V_i}$ forms an orthonormal basis for the space of class functions!!!

This, for example, tells us that there are $|\frac{G}{G}|$ many irreducible representations.

This also means that we can easily compute the irreducibles in a rep $V$: $(\chi_V, \chi_{V_i})$ is the number of $V_i$ in $V$.

Given a finite group $G$, we can write down a table, with columns indexed by conjugacy classes of $G$ and rows indexed by irreducible representations. The valued at $(i,j)$ is $\chi_{V_i}(g_j)$. The orthogonal condition exists for both rows and columns (for rows need to multiply by the size of the conjugacy class). This can be used to find the characters of the irreducible reps.




\subsection{Galois Theory}

Galois theory is very similar to covering space theory, however in this case the (absolute) Galois group is profinite, so a bit harder to understand. 

\subsubsection{Normality and Separability}


Let $F$ be a field. We will consider finite field (algebraic) extensions over $k$. Examples of them are  $F[x]/f(x)$ where $f$ is an irreducible polynomial over $k$. 
\begin{remark}
We will assume that we have an algebriac closure $\overline{F}$ over $F$. Note that this choice is not canonical and it is like choosing a universal cover together with a base-point, equivalently, choosing a point of $BG$, where $BG$ is viewed as a non-based space. Making a choice will trivialize a lot of other things. This probably will not play a big factor in the quals but it is something really important to keep in mind.
\end{remark}

Given an algebraic field extension $K/F$ and $k \in K$, then there exists an unique monic minimal polynomial $f \in F[x]$ such that $f(k) = 0$ and all other polynomial that does has $f$ as a factor (hence minimality). This is because we have a map $F[x] \rightarrow K$ sending $x \mapsto k$. Since the extension if algebraic, this has a nonzero ideal as a kernel, as $F[x]$ is a $PID$, this is generated by some monic polynomial $f$.

Here's the notion of normality:
\begin{definition}
Let $K/F$ be a finite extension, then it is normal if for every $k \in K$, the minimal polynomial $f_k(x) \in F[x]$ of $k$ splits completely over $K$, that is, we can find all the roots of $f_k$ in $K$. Equivalently, it is normal if for every irreducible $f \in F[x]$ which has a root in $K$, then it splits completely over $K$.
\end{definition}

Here's a criterion for normality: 
\begin{lemma}
Let $f \in F[x]$ be an irreducible polynomial, the field extension $K \coloneqq  F[x]/f(x)\ /F $ is normal iff it contains all the roots of $f$ ($f$ splits completely).
\end{lemma}

Here's some other equivalent notion for normality:
\begin{enumerate}
    \item All embeddings $K \xhookrightarrow{} \overline{F}$
    has the same image
    \item The automorphism group $Aut_F(K)$ acts transitively on the set of homomorphisms $K \xhookrightarrow{} \overline{F}$.
\end{enumerate}

Given an extension $K/F$, easy to see that there is a normal closure $N(K)/F$, where $N(K)$ is the mininal normal extension of including $K$ over $F$. It is generated by all the other roots of the minimal polynomial of elements of $K$. 

Given $f \in F[x]$ not necessarily irreducible, we have the splitting field of $F$ which is the smallest normal extension that splits $f$ completely.

\begin{example}
Examples of normal extensions are cyclotomic extensions ($\mathbb{Q}[\zeta_i]$ where $\zeta_i$ is a primitive $i$-th root of unity. Equivalently, it is the splitting field of $x^n -1$ in $\mathbb{Q}$.) 

Another one is $\mathbb{Q}[x]/(x^2 - 2)$, aka adjointing $\sqrt{2}$. In fact, any degree 2 extensions are normal. 

An example of a non-normal extension is $\mathbb{Q}[\sqrt{2}^3]$, because this is contained in $\mathbb{R}$ (really an embedding of this into $\mathbb{C}$), but the other two are complex. 
\end{example}

We also have the notion of separable:


\begin{definition}
A polynomial $f \in F[x]$ is separable if it has no repeated roots (in $\overline{F}$).
An extension $K/F$ is separable if any minimal polynomial of an element of $K$ is separable. 

An extension $K/F$ is purely-nonseperable if any minimal polynomial of an element of $K - F$ is non-separable.
\end{definition}

We have criterion for a polynomial to be separable:
\begin{lemma}
The following condition is equivalent:
\begin{enumerate}
    \item $f$ is separable
    \item $(f, f') = 1$, where $f'$ is the derivative of $f$.
    \item The discriminant of $f$ is 0, where discriminant is a homogeneous degree $2n-2$ polynomial in coefficients of $f$. It is something like $\prod_{i \neq j} (r_i - r_j)$, where $r_i, r_j$ are the roots of $f$.
\end{enumerate}
\end{lemma}

Here's an equivalent definition for seperable extension:

\begin{lemma}
An extension is separable iff there are exactly $[K:F]$ embedding from $K$ in $\overline{F}$ (or the normal closure of $F$.
\end{lemma}

Here's discriminant in low dimensions: in degree 2, the discriminant of $f = ax^2 + bx + c$ is $b^2 - 4ac$. In degree 3, the discriminant of $f = x^3 + px + q$ is $-4 p^3 - 27 q^2$ (note that this is not degree 4 polynomial as there are missing coefficients of $x^3$, which is set to 1.

For an irreducible polynomial $f$ to be nonseparable, we must have $f' = 0$. This can happen iff we are in char $p$ and $f$ is have entirely $x^{np}$ coefficients. 

\begin{example}
All char $0$ irreducible polynomials are separable. 

All char 0 field extensions and finite field extensions are separable. For a char $p$ field, any purely-nonseparable extensions are coming from $X^{p^n} - a$ for $a \in F$. 
Typical example is to take $F = \mathbb{F}_p (t)$ and $K = \mathbb{F}_p (t^{\frac{1}{p}})$
\end{example}

\begin{definition}
A field $F$ is called perfect if $\overline{F}/F$ is separable.
\end{definition}

Examples are char 0 fields and finite fields.


\subsubsection{Galois theory}
We first start with the definition of a Galois extension:

\begin{definition}
A field extension $K/F$ is Galois iff it is normal and separable. In this case, the Galois group $Gal(K/F)$ is defined as $Aut_K(F)$.
\end{definition}

Here's where the Galois group comes in:
\begin{lemma}
$|Gal(K/F)| = [K: F]$.
\end{lemma}

\begin{proof}
By normality, the Galois group acts transitively on the set of embedding of $K$ into $\overline{F}$. One can show that the action if also free. Separability implies the size of the set of embedding of $K$ into $\overline{F}$ is $[K:F]$.
\end{proof}

Normality extension composes, and we also have the following lemma:
\begin{lemma}
If we have $E/K/F$, and $E/F$ is normal (separable, Galois), then so is $E/K$.
\end{lemma}

Given a Galois extension $E/F$, we have the associated group $G = Gal(E/F)$. We have the following construction:

For any intermediate field extension $E/K/F$, then we have the associated group $Gal(E/K) \subset G = Gal(E/F)$. Conversely, given a subgroup $H$, then we can construct the subfield $E^H$ of the fixed points of $E$ under the $H$ action. 

Easily to see that $Gal(E/E^H) = H$. and $E^{Gal(E/K} = K$. Thus we have the following:

\begin{theorem}[Fundamental theorem of Galois Theory]
There exists an equivalence of category between the poset of intermediate extensions between $E/F$ and the poset of subgroup of $G = Gal(E/F)$.
\end{theorem}

\begin{remark}
This is completely analogous to the covering space with a base point situation.
\end{remark}

However, as with covering space, it is more interesting to think about extensions of $F$ not necessarily with an embedding, but in this case it gets a bit harder (there are profinite things for examples). Nevertheless we still have that $Aut (K/F) = N(Gal (E/K))/Gal (E/K)$. $N$ is the normalizer of $Gal(E/K)$ in $Gal(E/F)$. This means that if $K/F$ is Galois, then we have $Gal (K/F) = Gal(E/F) / Gal(E/K)$


\todo[inline]{Review the basics of field extensions and Galois theory, discriminants. How to know if a Galois extension is in $A_n$ etc. Examples of basic construction of Galois extensions that are like $\mathbb{Z}/2\mathbb{Z} \times \mathbb{Z}/2 \mathbb{Z}$ etc.}

\subsubsection{Galois theory of Cyclotomic Extensions}
The $n$-th cyclotomic field is the splitting field of $x^n-1$ over $\mathbb{Q}$. Note that $x^n -1$ is ofc not irreducible. It breaks into cyclotomic polynomials: 
$\Phi_n = \prod_{i} (x - \zeta_i)$ where $\zeta_i$ are the primitive $n$-th roots (there are $\phi(n)$ of them, where $\phi$ is ofc the Euler function). They are actually integer valued polynomials. We have that 
$$
x^n - 1 = \prod_{d | n} \Phi_d
$$

Thus we see that the $\mathbb{Q}[\zeta]$, where $\zeta$ is a primitive $n$-th root of unity,  is a degree $\phi(n)$ extension over $\mathbb{Q}$. Easy to see that this is the splitting field of $x^n -1$ as all other roots are powers of $\zeta$. The Galois group is $(\mathbb{Z}/n\mathbb{Z})^\times \cong \mathbb{Z}/\phi(n)\mathbb{Z}$, acting on $\zeta$ as $\zeta \mapsto \zeta^i$ for $i$ coprime to $n$. Ofc this determined where all other powers of $\zeta$ goes to.

\todo[inline]{Primitive element theorem and normal element theorem.}
\subsubsection{Primitive Element and Normal Basis Theorems}

Here's two nice theorems:
\begin{theorem}[Primitive element theorem]
For any separable field extension $K/F$, there exists an element of $k \in K$ such that $K = F(k)$.
\end{theorem}

\begin{remark}
The counterexample is the degree $p^2$ extension $\mathbb{F}_p(U^\frac{1}{p}, T^\frac{1}{p})$ over $\mathbb{F}_p(U,T)$, any rational polynomial on $U^\frac{1}{p}, T^\frac{1}{p}$ has $p$-th power in $\mathbb{F}^p(U,T)$.
\end{remark}

Here's the normal basis theorem:
\begin{theorem}[Normal basis theorem]
For any Galois extension $K/F$ with Galois group $G$, there exists an element $\sigma$ such that $G \sigma$ spans $K$ over $F$. Equivalently, we have $F[G] \cong K$ as left $F[G]$ modules. 
\end{theorem}

\begin{remark}
Of course when the extension is not normal this is not true, as the size of the automorphism group is smaller the $[K:F]$.
\end{remark}

\subsubsection{Galois Theory of Finite Fields}
\todo[inline]{Just all the standard stuff, the groups are cyclic generated by the Frobenius element... All group of units is cyclic...}


\subsubsection{Examples of Galois Extensions with a given Galois Group}



\subsection{Ring Theory and Algebraic Number Theory}
\todo[inline]{Write about integrally closed, UFD etc. Prime factorization in a Dedekind domain, quadratic extensions $\mathbb{Z}[\sqrt{n}]$ and rather is it is integrally closed/UFD.}


\subsection{Linear Algebra}


Given a finite dimensional vector space $V$ over a field $k$ and $T: V \rightarrow T$, then we want to understand how to decompose $T$. The idea is simple: use algebraic geometry: consider $T$ as $x$ in $k[x]$, then we can view $V$ as a coherent sheave over $\mathbb{A}^1 = spec k[x]$. Being finite dimensional, $V$ is supported at finite points. It is supported on the closed sub-scheme cut out by a polynomial $f$ on $k[x]$. This $f$ is the mininal polynomial on $x$ such that $f(T) = 0$. This is called the minimal polynomial of $T$: 

\begin{definition}
The minimal polynomial $f$ of $T$ is the unique lowest degree polynomial in $T$ such that $f(T) = 0$.
\end{definition}

On the other hand, we also have the characteritic polynomial, which is a degree $n$ polynomial in T:
\begin{definition}
$char(x) coloneqq det (xId - T)$ is called the characteritic polynomial in T
\end{definition}

It's $x^{n-1}$ coefficient is $-tr(T)$ and the $x^0$ coefficient is $(-1)^{n-1} \ det \ T$. 

The Cayley-Hamilton theorem says that 
\begin{theorem}[Cayley-Hamilton]
$char(T) = 0$
\end{theorem}
Thus the mininal polynomial factors the characteristic polynomial.

In fact, using the classification theory of finitely generated modules over a PID (in this case $k[x]$), we can do more:

Using the invariant factor decomposition, we see that $V \cong k[x]/(f_1) \oplus k[x]/(f_2) .. \oplus k[x]/(f_n)$ with $f_1 | f_2 | ... | f_n$. $f_n$ in this case will be the minimal polynomial. 

In linear algebra, this means that we can find direct summands $V_i$ for $V$ such that $T$ sends $V_i$ to $V_i$ ($T$ looks block diagonal).For each $V_i$, we have basis for the $T$ on all but last basis is off diagonal 1: just think about how $x$ acts on $k[x]/(f(x))$ with basis $1, x, x^2, ...$. This is the rational canonical form. The minimal polynomial of $T|V_i$ is ofc $f_i$, equaling the characteristic polynomial of $T|V_i$ (just by degree reasons). Note that the first basis vector $v$ has $v, T\ v, T^2 v $ spans $V_i$.

This off diagonal form is called the rational canonical form. Note that this exists for any field $k$. 

Now if $k$ is algebraically closed, then we know that we can factor each $f$ to $(x-a_1)^{e_1} (x - a_2)^{e_2}...$,

So we have the case now $T$ has characteristic polynomial (equals minimal polynomial) is $(x-a)^n$. This means that $T ' = T - a Id$ has characteristic polynomial $x^n$. There is a basis here $x^{n-1} = (T')^{n-1} 1,... x = T' 1, 1$, in this basis, $x^{n-1}$ is the eigenvector for $T'$ with eigenvalue $0$, thus it is an eigenvector for $T$ with eigenvalue $a$. Moreoever, the $i$-th basis vector gets annihilated by $(T')^i$ (this is all theory for nilpotent matricies). Thus we see that in this basis, $T'$ can be written as diagonal $0$, one up than diagonal $1$. Therefore $T$ can be written as diagonal $a$, one up that diagonal $1$. 

For a more general $T$, then we see that we can find a basis with blocks looking like such above. This is called the Jordan normal form of $T$.

\begin{remark}
Note that this is true for any matrix whose characteristic polynomial factors into linear polynomials. Ofc this is true for any matrix over algebraically closed fields.
\end{remark}


\end{document}